---
title: "SURVIVAL_ANALYSIS"
output: html_document
date: 20 March, 2019
---

https://stat.ethz.ch/R-manual/R-devel/library/survival/html/ovarian.html

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(kableExtra)
```

#### Cleaning data

```{r, include=TRUE}
#rx is treatment group
#residual disease present (1=no, 2=yes)
#ecog.ps: ECOG performance status (1 is better, see reference)

data("ovarian")
head(ovarian)

ovarian$rx <- factor(ovarian$rx,levels = c("1", "2"), labels = c("A", "B")) 
ovarian$resid.ds <- factor(ovarian$resid.ds, levels = c("1", "2"), labels = c("no", "yes")) 
ovarian$ecog.ps <- factor(ovarian$ecog.ps, levels = c("1", "2"), labels = c("good", "bad"))
head(ovarian)
```

#### Kaplan-Meier (survival package)

```{r, include=TRUE}
#futime: survival or censoring time
#fustat: censoring status

#1 create km with survfit for whole cohort
km_fit <- survfit(Surv(futime, fustat) ~ 1, data=ovarian) 
print(km_fit)
summary(km_fit)
#Use times() parameter to select some clinically relevant time points
summary(km_fit, times = c(90, 270, 450)) 


#2 create km with survfit for cohort according to presence of residual disease
km_grp_fit <- survfit(Surv(futime, fustat) ~ resid.ds, data=ovarian)
print(km_grp_fit)
summary(km_grp_fit)
```
-time: time points on the curve. n.risk: the number of subjects at risk at time t. n.event: the number of events that occurred at time t.lower,upper: lower and upper confidence limits for the curve, respectively



####plot KM curve in base R
```{r, include=TRUE}
#1
plot(km_fit, xlab="Days", main = 'KMPlot')

#2
plot(km_grp_fit, xlab="Days", main = 'KMPlot', conf.int=TRUE, col=c("red","blue"))
legend(100, .2, c("No Residual Disease", "Residual Disease"), lty = c(1:2), col=c("red","blue"))
```



#### plot KM curve with survminer package
```{r, include=TRUE}
library(survminer)
km_fit <- survfit(Surv(futime, fustat) ~ 1, data=ovarian) 
ggsurvplot(km_fit, data = ovarian , risk.table = TRUE, pval = TRUE, conf.int = TRUE)
km_grp_fit <- survfit(Surv(futime, fustat) ~ resid.ds, data=ovarian)
ggsurvplot(km_grp_fit, data = ovarian , risk.table = TRUE, pval = TRUE, conf.int = TRUE)
```

####Log-rank test
**to compare survival curves of two or more groups. Null hypothesis is that that survival curves of two populations do not differ**


```{r}
survdiff <- survdiff(Surv(futime, fustat) ~ resid.ds, data = ovarian)
survdiff
```
**Results:**p = 0.06. Don't have sufficient evidence to reject the null hypothesis.


####Cox Proportional hazards (cph) regression (multivariable survival analysis)
**to describe the effect of variables on survival**

```{r}
#create survival object
surv1 <- Surv(time = ovarian$futime, event = ovarian$fustat) 
surv1

#do regression
fit_cph <- coxph(surv1 ~ rx + resid.ds + age + ecog.ps, data = ovarian)
summary(fit_cph)
cox_fin<-round(exp(cbind(HR = coefficients(fit_cph), confint(fit_cph))), digits=2)

cox_fin %>%
  kable() %>%
  kable_styling(full_width = F)

#Interpretations:
#HR = 1: No effect
#HR < 1: Reduction in the hazard
#HR > 1: Increase in Hazard
```
-age is the only signif variable in this model, positively associated with outcome event (mortality)
-We see from our CPH regression results, age is positively associated with hazard/risk of death (HR = 1.13 (1.03, 1.24), p = 0.008), accounting for treatment, residual disease and ecog score
